---
title: "r4ds_communicate_exercises_27"
author: "Otis"
date: "10/28/2017"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

## Chapter 27

Import packages
(I'm using include = FALSE in the R script but it's not outputting any code in the rmd file)
```{r, include = FALSE}
library(tidyverse)
library(scales)
```


### 27.2.1

_1_  
Made a notebook  
_2_  
The knit function works (yay!)  
_3_  
Both have the similar workflow of allowing for markdown text with r-scripts being able to be run & output code throughout the file
RNotebook doesn't have the knit option functionality, or the as descriptive YAML header up top
From what I've seen so far they are quite similar
_4_  
Overall the knit plots to HTML, PDF & Word look very similar. Each has respective style differences for its file type (and is rendered in a different program accordingly). Things like font / color change slightly but overall structure stays the same

### 27.3.1

_1_  
#### Otis' CV
Otis Skipper  
otisskipper@gmail.com  
828.275.4007

##### Education
* University Of North Carolina Chapel Hill
 *  __2012-2016__  
 * B.S. In Mathematic
 * 3.9 GPA

#### Work Experience
* Red Ventures Data Analyst
  * __2016-Present__
  * Create reporting for business teams to make actionable insights from data
  * Validate that technologies are producing data accurately

_2_  
None of these seem to be working in rmarkdown output, but they are all from the rmarkdown cheatsheet
This is question 2.[^1]  
[^1]: Here is the footnote.

horizontal rule/slide break:  
***  
> block quote


_3_

Ran it in separate markdown file

### 27.4.7

_1_
```{r}
knitr::opts_chunk$set(
  echo = FALSE
)
```

```{r}
knitr::kable(diamonds %>% 
  group_by(cut) %>% 
  summarise(avg_carat = mean(carat)))

knitr::kable(diamonds %>% 
  group_by(color) %>% 
  summarise(avg_carat = mean(carat)))

knitr::kable(diamonds %>% 
  group_by(clarity) %>% 
  summarise(avg_carat = mean(carat)))

```
_2_  
Table describing the largest 20 diamonds
```{r}


knitr::kable(diamonds %>% 
  arrange(desc(carat)) %>% 
  top_n(20))
```

_3_

```{r}
knitr::kable(diamonds %>% 
  mutate(price = comma(price)) %>% 
  arrange(desc(carat)) %>% 
  top_n(20))
```

Percent of diaomonds > 2.5 carats ~= 0.2%
```{r}
total <- diamonds %>%
  count()
over <- diamonds %>% 
  filter(carat > 2.5) %>% 
  count()
pct <- over / total
paste(pct)
```

_4_
Run it & d = 11, all timestamps are equal.  
Change value of b & run again: value of d stays the same (so do timestamps for a, b, c). Only value of b & nowb change
```{r, cache = TRUE, include = TRUE, eval = TRUE}
a <- 2
nowa <- lubridate::now()
a
nowa
```

```{r, cache = TRUE}
b <- 2 + a
nowb <- lubridate::now()
b
nowb
```

```{r, cache = TRUE}
c <- 3 + a
nowc <- lubridate::now()
c
nowc
```

```{r, cache = TRUE}
d <- a + b + c
nowd <- lubridate::now()
d
nowd
```

