---
title: "R4DS Exercises 14-15"
output: html_notebook
---

#### Load Packages
Code to load necessary packages
```{r}
library(tidyverse)
library(stringr)
library(forcats)
```

# Chapter 14

####__14.2.5__  

_1_  
```{r}
?paste
?paste0
paste0(1:12, c("st", NA, "rd", rep("th", 9)))
str_c(1:12, c("st", NA, "rd", rep("th", 9)))
```
paste0 has a default separator of nothing ("") - just like calling paste(..., sep = "")  
Functions are equivalent to str_c in the stringr package  
Paste functions will combine a legitimate string & NA into a string, whereas Stringr overwrites the combination with NA  
ie: differences in 2nd value of code output from above  


_2_   
sep is used to either combine strings explicitly passed to str_c or strings across vectors passed to str_c  
collapse is used to combine strings within a vector


_3_

```{r}
x <- "abcde"
len <- str_length(x)
str_sub(x, len/2+1, len/2+1)
```
If string length is even this function grabs the character at the ceiling(length) position

_4_ 
```{r}
?str_wrap
output <- "str_wrap is good for formatting strings into human readable scentences & paragraphs and adjusting things like width/indenting. You may want to use it when outputting long text strings into the consolse."
cat(str_wrap(output, width = 30), "\n")

```

_5_  
str_trim will trim whitespace from the beginning/end of a string
str_pad will add whitespace to beginning/end

_6_  
Create function to deal with comma separated strings
```{r}

oxfordComma <- function(v) {
if(length(v) == 0){
  print("Length 0 vector - you'll get a 0 character vector")
}
commas <- str_c(v[1:length(v)-1], collapse = ", ")
output <- str_c(commas, v[length(v)], sep = ", and ")
output
}
x <- c("a", "b", "c", "d")
oxfordComma(x)

```


####__14.3.1.1__  
_1_  
Explanation comes from earlier in book:  
You need "\\\\" to just match "\" (which has to be written "\\" as a string)
"\" - being used as escape but nothing follows it
"\\" - escape into a single "\", but actually have to match "\\"
"\\\" - escapte into a single "\" but then escape into nothing after that

_2_ 
```{r}
str_view("\"'\\", "\"'\\\\" )
```

_3_
```{r}
str_view(".x.y.z", "\\..\\..\\.." )
```


####__14.3.2.1__  
_1_  
```{r}
str_view("$^$", "\\$\\^\\$")
```

_2_  
```{r}
str_view(stringr::words, "^y", match = TRUE)
str_view(stringr::words, "x$", match = TRUE)
str_view(stringr::words, "^...$", match = TRUE)
str_view(stringr::words, "^.......", match = TRUE)

```

####__14.3.3.1__  
_1_  
```{r}
str_view(stringr::words, "^[aeiou]", match = TRUE)
str_view(stringr::words, "^[^aeiou]+$", match = TRUE)
str_view(stringr::words, "[^e]ed$" , match = TRUE)
str_view(stringr::words, "(ing|ise)$" , match = TRUE)
```

_2_
```{r}
str_view(stringr::words, "[^c]ie" , match = TRUE)

```

_3_  
In dataset q is always followed by u
```{r}
str_view(stringr::words, "q[^u]", match = TRUE)
```

_4_  
Assume british english uses 'our' instead of 'or' (favor, color, etc...)
```{r}
str_view(stringr::words, "our", match = TRUE)

```

_5_
```{r}
str_view("828-275-4007", "\\d\\d\\d-\\d\\d\\d-\\d\\d\\d\\d", match = TRUE)
```


####__14.4.2__  
 
_1_
```{r}
#1
str_view(stringr::words, "(^x)|(x$)", match = TRUE)

word1 <- data.frame(words)
word1$endx <- str_detect(words, "x$")
word1$startx <- str_detect(words, "^x")
word1 %>% 
  filter(endx == TRUE | startx == TRUE)

#2
str_view(stringr::words, "^[aeiou].*[^aeiou]$", match = TRUE)

word2 <- data.frame(words)
word2$startVowel <- str_detect(words, "^[aeiou]")
word2$endCons <- str_detect(words, "[^aeiou]$")
word2 %>% 
  filter(startVowel == TRUE & endCons == TRUE)


#3

# Don't know the order the vowels will appear
# Can guaruntee 5 vowels - but not that they're unique amongst vowels
str_view(stringr::words, ".*[aeiou].*[aeiou].*[aeiou].*[aeiou].*", match = TRUE)
# For a single Regex have to have every possible permutation: 5! = 120 ()
# Found a permutation of all 120 combos on stack overflow (https://unix.stackexchange.com/questions/156134/grep-to-find-words-with-all-vowels)
str_view("eiousa", "a.*e.*i.*o.*u|a.*e.*i.*u.*o|a.*e.*o.*i.*u|a.*e.*o.*u.*i|a.*e.*u.*i.*o|a.*e.*u.*o.*i|a.*i.*e.*o.*u|a.*i.*e.*u.*o|a.*i.*o.*e.*u|a.*i.*o.*u.*e|a.*i.*u.*e.*o|a.*i.*u.*o.*e|a.*o.*e.*i.*u|a.*o.*e.*u.*i|a.*o.*i.*e.*u|a.*o.*i.*u.*e|a.*o.*u.*e.*i|a.*o.*u.*i.*e|a.*u.*e.*i.*o|a.*u.*e.*o.*i|a.*u.*i.*e.*o|a.*u.*i.*o.*e|a.*u.*o.*e.*i|a.*u.*o.*i.*e|e.*a.*i.*o.*u|e.*a.*i.*u.*o|e.*a.*o.*i.*u|e.*a.*o.*u.*i|e.*a.*u.*i.*o|e.*a.*u.*o.*i|e.*i.*a.*o.*u|e.*i.*a.*u.*o|e.*i.*o.*a.*u|e.*i.*o.*u.*a|e.*i.*u.*a.*o|e.*i.*u.*o.*a|e.*o.*a.*i.*u|e.*o.*a.*u.*i|e.*o.*i.*a.*u|e.*o.*i.*u.*a|e.*o.*u.*a.*i|e.*o.*u.*i.*a|e.*u.*a.*i.*o|e.*u.*a.*o.*i|e.*u.*i.*a.*o|e.*u.*i.*o.*a|e.*u.*o.*a.*i|e.*u.*o.*i.*a|i.*a.*e.*o.*u|i.*a.*e.*u.*o|i.*a.*o.*e.*u|i.*a.*o.*u.*e|i.*a.*u.*e.*o|i.*a.*u.*o.*e|i.*e.*a.*o.*u|i.*e.*a.*u.*o|i.*e.*o.*a.*u|i.*e.*o.*u.*a|i.*e.*u.*a.*o|i.*e.*u.*o.*a|i.*o.*a.*e.*u|i.*o.*a.*u.*e|i.*o.*e.*a.*u|i.*o.*e.*u.*a|i.*o.*u.*a.*e|i.*o.*u.*e.*a|i.*u.*a.*e.*o|i.*u.*a.*o.*e|i.*u.*e.*a.*o|i.*u.*e.*o.*a|i.*u.*o.*a.*e|i.*u.*o.*e.*a|o.*a.*e.*i.*u|o.*a.*e.*u.*i|o.*a.*i.*e.*u|o.*a.*i.*u.*e|o.*a.*u.*e.*i|o.*a.*u.*i.*e|o.*e.*a.*i.*u|o.*e.*a.*u.*i|o.*e.*i.*a.*u|o.*e.*i.*u.*a|o.*e.*u.*a.*i|o.*e.*u.*i.*a|o.*i.*a.*e.*u|o.*i.*a.*u.*e|o.*i.*e.*a.*u|o.*i.*e.*u.*a|o.*i.*u.*a.*e|o.*i.*u.*e.*a|o.*u.*a.*e.*i|o.*u.*a.*i.*e|o.*u.*e.*a.*i|o.*u.*e.*i.*a|o.*u.*i.*a.*e|o.*u.*i.*e.*a|u.*a.*e.*i.*o|u.*a.*e.*o.*i|u.*a.*i.*e.*o|u.*a.*i.*o.*e|u.*a.*o.*e.*i|u.*a.*o.*i.*e|u.*e.*a.*i.*o|u.*e.*a.*o.*i|u.*e.*i.*a.*o|u.*e.*i.*o.*a|u.*e.*o.*a.*i|u.*e.*o.*i.*a|u.*i.*a.*e.*o|u.*i.*a.*o.*e|u.*i.*e.*a.*o|u.*i.*e.*o.*a|u.*i.*o.*a.*e|u.*i.*o.*e.*a|u.*o.*a.*e.*i|u.*o.*a.*i.*e|u.*o.*e.*a.*i|u.*o.*e.*i.*a|u.*o.*i.*a.*e|u.*o.*i.*e.*a")

#str_detect helps simplify
word3 <- data.frame(words)
word3$a <- str_detect(words, "a")
word3$e <- str_detect(words, "e")
word3$i <- str_detect(words, "i")
word3$o <- str_detect(words, "o")
word3$u <- str_detect(words, "u")
word3 %>% 
  filter(a == TRUE & e == TRUE & i == TRUE & o == TRUE & u == TRUE)


```
No words in the stringr::words data with all vowels

_2_

```{r}

word4 <- data.frame(words)
word4$length <- str_length(word4$words)
word4$vowels <- str_count(word4$words, "[aeiou]")
word4 %>% 
  mutate(vowelPct = vowels / length) %>% 
  arrange(desc(vowelPct))
word4 %>% 
  arrange(desc(vowels))

```
"appropriate", "associate", "available", "colleague", "encourage", "experience", "individual", "television" all have 5 vowels (all tied for highest count)  
"a" has the highest vowel pct (100%), followed by "area" & "idea" at 75%

####__14.4.6.1__  
_1_
```{r}
"apples, pears, and bananas" %>% 
  str_replace("and", "") %>% 
  str_split(",")
```

_2_
If a string has multiple spaces than str_split(" ") will bring in blank spaces as strings

_3_
```{r}
str_split("test string", "")
```
Breaks the string into all individual characters

####__15.4.1__  
 _1_
There are 41 values for tvhour above 18 (which means people are only wathcing tv then sleeping - some values are even 24 hours which means no sleep)
Should consider excluding outliers or unreasonable values (then mean could be a good value)

```{r}
gss_cat %>% 
  select(tvhours) %>% 
  filter(tvhours > 18) %>% 
  count()
 
```

_2_
Year: principled  
Marital: arbitrary  
Age: principled  
Race: arbitrary  
rincome: principled  
partyid: arbitrary  
relig: arbitrary  
denom: arbitrary  
tvhours: principled

_3_  
The levels in the front get plotted as the bottom values (No answer and Don't know were the other 2 levels in the front)


####__15.5.1__  
Relatively flat trends over time looking at the bar graph below
Small trend of more Independents & 'others' but no large trends in any direction
 _1_
```{r}
df <- gss_cat %>%
  mutate(partyid = fct_collapse(partyid,
    other = c("No answer", "Don't know", "Other party"),
    rep = c("Strong republican", "Not str republican"),
    ind = c("Ind,near rep", "Independent", "Ind,near dem"),
    dem = c("Not str democrat", "Strong democrat")
  ))

df %>% 
  ggplot(aes(x = year, group = partyid, fill = partyid)) + 
  geom_bar(position = "fill")  
```
_2_
Collapse rincome into smaller groups
```{r}
gss_cat %>%
  mutate(rincome = fct_collapse(rincome,
    unkown = c("No answer", "Don't know", "Refused", "Not applicable"),
    ZeroTo5K = c("Lt $1000", "$1000 to 2999", "$3000 to 3999", "$4000 to 4999"),
    FiveKto10K = c("$5000 to 5999", "$6000 to 6999", "$7000 to 7999", "$8000 to 9999"),
    TenKOrMore = c("$10000 - 14999", "$15000 - 19999", "$20000 - 24999", "$25000 or more")
  ))
```

