---
title: "Accelerator Exercises Ch. 28"
author: "Otis"
date: "11/1/2017"
output: html_document
---

### Load Packages
```{r, message = FALSE}
library(tidyverse)
```


### Chapter 28

#### 28.2.1
_1_    
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    title = "Here's a custom title for my plot",
    subtitle = "And here's a custom subtitle",
    colour = "Car type"
  )
```

_2_   
Break out geom smooth function by class to see how displ and hwy behave within the class of each car type.
For car types with a small number of data points the geom_smooth function doesn't have enough points for a strong fit, so we ignore the error message
```{r, warning = FALSE}
ggplot(mpg,aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  facet_grid(.~class)+
  geom_smooth()
```

_3_  
```{r}
diamonds %>% 
  ggplot(aes(price))+
  geom_histogram(bins = 20) +
  labs(
    title = "Historgram Distribution of Diamond Prices",
    caption = "data taken from diamonds data set within ggplot",
    x = "price of diamond",
    y = "count of diamonds"
  ) +
  theme(legend.position = "none")
```


#### 28.3.1

_1_  

First Solution I got: There's a more elegant way to do this but it technically works
```{r}

displ <- c(Inf, Inf, -Inf, -Inf)
hwy <- c(Inf, -Inf, -Inf, Inf)
txt <- c("top right", "bottom right", "bottom left", "top left")
tmp <- tibble(displ, hwy, txt)
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +

  geom_text(aes(label = txt), data = tmp, hjust ="top", vjust = "right") +
  geom_text(aes(label = txt), data = tmp, hjust ="bottom", vjust = "left") +
  geom_text(aes(label = txt), data = tmp, hjust ="bottom", vjust = "right") +
  geom_text(aes(label = txt), data = tmp, hjust ="top", vjust = "left")
 
```

Better solution (suggestion from Chris S. after meeting in Breakout Session)
```{r}
label_tr <- tibble(displ = Inf, hwy = Inf)
label_br <- tibble(displ = Inf, hwy = -Inf)
label_tl <- tibble(displ = -Inf, hwy = Inf)
label_bl <- tibble(displ = -Inf, hwy = -Inf)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes (label = "topr"), data = label_tr, vjust = "top", hjust = "right") +
    geom_text(aes (label = "botr"), data = label_br, vjust = "bottom", hjust = "right") +
  geom_text(aes (label = "topl"), data = label_tl, vjust = "top", hjust = "left") +
  geom_text(aes (label = "botl"), data = label_bl, vjust = "bottom", hjust = "left") 

```


_2_  
```{r}
?annotate
```
Can pass in small annotations (Such as text labels) as a vector instead of having to put them into a data frame

_3_   

If text is generically defined across graph then it will pass through to each facet
```{r}

label <- tibble(
  displ = Inf,
  hwy = Inf,
  label = "Increasing engine size is \nrelated to decreasing fuel economy."
)

ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_text(aes(label = label), data = label, vjust = "top", hjust = "right") +
  facet_grid(drv~.)
```

But if text is restricted to a certain data point (as in the example below) that our data is being facted on, then it will only translate to the specific facet that has that value. ie: corvette is associated with drv r and thus only shows up in that specific facet
```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class) +
  facet_grid(drv~.)
```

_4_  
"The `fill` aesthetic controls the background colour of the label" - R documentation

_5_  
arugments of arrow:
angle, length, ends, type

Examples of different uses for arrows
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_segment(aes(x=displ, y=hwy, xend = displ +1, yend = hwy + 1), arrow = arrow(15, length = unit(0.1, "inches"), ends = "last", type = "closed"))

  
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  geom_segment(aes(x=2, y=40, xend = 7, yend = 15), arrow = arrow(40, length = unit(0.2, "inches"), ends = "both", type = "open"))
```


#### 28.4.4
_1_  
Load Hexbin Package
```{r, message = FALSE}
library(hexbin)
```

Original code used scale_colour_gradient - need to use scale_fill_gradient to get red/white values to appear

```{r}

df <- tibble(
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() +
  scale_fill_gradient(low = "white", high = "red") +
  coord_fixed()
```

_2_

First argument to both functions is `...`. For labs this is "A list of new name-value pairs", for scale this is "Other arguments passed on to continuous_scale to control name, limits, breaks, labels and so forth." They are both generic dots arguments, but control different things for each function

_3_
```{r}
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue")) +
  scale_y_continuous("President Number", breaks = seq(34,44), label = seq(34,44)) +
  scale_x_date("Year", date_breaks = "4 year", date_labels = "'%y") +
  ggtitle("Presidential Terms By Party") +
  geom_text(aes(label = presidential$name), nudge_y = -.4, nudge_x = 2100, show.legend =FALSE)
?geom_text

```

_4_

```{r}
ggplot(diamonds, aes(carat, price)) +
  geom_point(aes(colour = cut), alpha = 1/20) +
  guides(colour = guide_legend(override.aes = list(alpha=1)))
```

